Indel Historian: phylogenetic ancestral reconstruction

### 2016-07-22 doc/main.tex
refs

Successful techniques include using Bayesian decision theory to maximize the expected
score under the metrics \cite{TCOFFEE,ProbCons,AMAP,FSA}
and fine-tuning the alignment scoring function to reflect known details of
protein sequences under selection for folding, such as the reduced likelihood of indels in
hydrophobic regions \cite{MAFFT,MUSCLE}

These have been accompanied by performance optimizations at the rate-limiting steps
of multiple alignment, including all-versus-all sequence comparison \cite{FSA,MUSCLE}

@@

Previous work showed that, for the purposes of estimating
molecular evolutionary parameters
such as the rates of insertion-deletion events \cite{Westesson2012-zg}
or the ratio between synonymous abd nonsynonymous substitution rates \cite{Redelings},
it is best to use an explicit probabilistic model of the sequence evolution process
and to treat sequence alignment rigorously as a missing data estimation problem
under this model

@@

A previously published benchmark of alignment methods on simulated sequence data
showed that most alignment algorithms introduce systematic biases into the
estimation of indel rates, and that these biases are typically substantially worse
at higher indel rates.

@@

We also report the results of benchmarking ProtPal on the BAliBase 3.0 benchmark

@@

\cite{Westesson2012-zg}

@@

Comparison to MCMC tools
BaliPhy, PhylogenyCafe, HandAlign

@@

Decision theory may also be useful in conjunction with MCMC tools \cite{HermanHeinEtc}

>>>>

Successful techniques include using Bayesian decision theory to maximize the expected
score under the metrics \citep{NotredameEtAl2000,DoEtAl2005,SchwartzPachter2007,BradleyEtAl2009}
and fine-tuning the alignment scoring function to reflect known details of
protein sequences under selection for folding, such as the reduced likelihood of indels in
hydrophobic regions \citep{KatohEtAl2005,Edgar2004b,LarkinEtAl2007}

PRANK\citep{LoytynojaGoldman2005}, Muscle \citep{Edgar2004b}, ClustalW\citep{LarkinEtAl2007}, MAFFT\citep{KatohEtAl2005}

These have been accompanied by performance optimizations at the rate-limiting steps
of multiple alignment, including all-versus-all sequence comparison \citep{BradleyEtAl2009,Edgar2004b}

@@

Previous work showed that, for the purposes of estimating
molecular evolutionary parameters---such as the rates of insertion-deletion events \citep{Westesson2012-zg},
the ratio between synonymous and nonsynonymous substitution rates \citep{Redelings2014},
or other evolutionary analyses \citep{LoytynojaGoldman2008}---it is best to use an explicit probabilistic model of the sequence evolution process
and to treat sequence alignment rigorously as a missing data estimation problem
under this model

@@

A previously published benchmark of alignment methods on simulated sequence data
showed that most alignment algorithms introduce systematic biases into the
estimation of indel rates, and that these biases are typically substantially worse
at higher indel rates \citep{Westesson2012-zg}.

@@

We also report an assessment of the structural alignment accuracy of IndelHistorian on the benchmark datasets BAliBase 3.0 \citep{ThompsonEtAl2005}
and PREFAB 4.0 \citep{Edgar2010}

@@

\citep{Westesson2012-zg}

@@

Comparison to MCMC tools
BaliPhy \citep{RedelingsSuchard2005,RedelingsSuchard2007,Redelings2014},
StatAlign \citep{NovakEtAl2008,HermanEtAl2014},
HandAlign \citep{WestessonBarquistHolmes2012}

@@

Decision theory may also be useful to produce consensus alignments summarizing an MCMC run \citep{HermanEtAl2015}
